### 2013-03-10 doc/sosp13/memory.tex
lots of edits. more coming

\paragraph{Global memory addressing} Grappa provides two methods for
storing data in the global memory. The first is a distributed heap
striped across all the machines in the system in a block cyclic fashion.
The \texttt{global\_malloc} and \texttt{global\_free} calls are used to
allocate and deallocate memory in the global heap.  Addresses to memory
in the global heap use \textbf{linear addresses}.  Choosing the block
size involves trading off sequential bandwidth against aggregate random
access bandwidth. Smaller block sizes help spread data across all the
memory controllers in the cluster, but larger block sizes allow the
locality-optimized memory controllers to provide increased sequential
bandwidth. The block size, which is configurable, is typically set to 64
bytes, or the size of a single hardware cache line, in order to exploit
spatial locality when available. The heap metadata is stored on a single
node. Currently all heap operations serialize through this node; while
this has been sufficient for our benchmarks, in the future Grappa will
provide parallel performance through combining~\cite{MAMA,flatcombining}.

Grappa also allows any local data on a core's stacks or heap to be
exported to the global address space to be made accessible to other
cores across the system. Addresses to global memory allocated in this
way use \textbf{2D global addresses}.  This uses a traditional PGAS
addressing model, where each address is a tuple of a rank in the job (or
global process ID) and an address in that process. The lower 48 bits of
the address hold a virtual address in the process. The top bit is set to
indicates that the reference is a 2D address (as opposed to linear
address). This leaves 15 bits for network endpoint ID, which limits our
scalability to $2^{15}$ endpoints. Any node-local data can be made
accessible by other nodes in the system by wrapping the address and node
ID into a 2D global address. This address can then be accessed with a
delegate and can also be cached by other nodes. At the destination the
address is converted into a canonical x86 address by replacing the upper
bits with the sign-extended upper bit of the virtual address. 2D
addresses may refer to memory allocated from a single processes' heap or
from a task's stack. Figure~\ref{fig:memory-structure} shows how 2D and
linear addresses can refer to other cores' memory.

@@

\comment{
    \begin{description}\small
      \item[ \texttt{ global\_address global\_malloc( size )} ] \hfill \\
      \item[ \texttt{ global\_free( global\_address )} ] \hfill \\
        Allocates and frees memory in the global heap
      \item[ \texttt{ delegate\_read( global\_address, local\_var )} ] 
      \item[ \texttt{ delegate\_write( global\_address, local\_var )} ] 
      \item[ \texttt{ delegate\_cas( global\_address, local\_var )} ] 
      \item[ \texttt{ delegate\_fetch\_inc( global\_address, local\_var )} ] 
\hfill \\
        Performs a memory operation at the home core of a global address
      \item[ \texttt{ cache\_acquire( global\_address, local\_buf, \{RO,RW,WO\})} ]
      \item[ \texttt{ cache\_release( global\_address, local\_buf )} ] 
\hfill \\
Perform cache operations to acquire/release global data.  Acquire, returns a local pointer after all data has been copied to the local node.  Release, optionally writes data back to global memory and frees up management resources.
    \end{description}
}

>>>>

\paragraph{Global memory addressing} Grappa provides two methods for storing
data in the global memory. The first is a distributed heap striped across all
the machines in the system in a block cyclic fashion. The
\texttt{global\_malloc} and \texttt{global\_free} calls are used to allocate
and deallocate memory in the global heap. Addresses to memory in the global
heap use \textbf{linear addresses}. Choosing the block size involves trading
off sequential bandwidth against aggregate random access bandwidth. Smaller
block sizes help spread data across all the memory controllers in the cluster,
but larger block sizes allow the locality-optimized memory controllers to
provide increased sequential bandwidth. The block size, which is configurable,
is typically set to 64 bytes, or the size of a single hardware cache line, in
order to exploit spatial locality when available.

%%% is the below important? -Luis
\comment{The heap metadata is stored on a single node. Currently all heap
allocation and free operations are serialized through this node; while this
has been sufficient for our benchmarks, it is easy to relax in the future
Grappa will provide parallel performance through
combining~\cite{MAMA,flatcombining}.}

Grappa also allows any local data on a core's stacks or heap to be
exported to the global address space to be made accessible to other
cores across the system. Addresses to global memory allocated in this
way use \textbf{2D global addresses}.  This uses a traditional PGAS
addressing model, where each address is a tuple of a rank in the job (or
global process ID) and an address in that process. The lower 48 bits of
the address hold a virtual address in the process. The top bit is set to
indicates that the reference is a 2D address (as opposed to linear
address). This leaves 15 bits for network endpoint ID, which limits our
scalability to $2^{15}$ endpoints. Any node-local data can be made
accessible by other nodes in the system by wrapping the address and node
ID into a 2D global address. This address can then be accessed with a
delegate operation and even be cached by other nodes. The
address is converted at the destination into a canonical x86 address by replacing the upper
bits with the sign-extended upper bit of the virtual address. 2D
addresses may refer to memory allocated from a single processes' heap or
from a task's stack. Figure~\ref{fig:memory-structure} shows how 2D and
linear addresses can refer to other cores' memory.

@@
